---
title: "The Tampa Bay Climate Chronicle"
subtitle: "A 10-Year Meteorological Investigation & Extreme Weather Dispatch (2015–2024)"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    css: styles.css
    include-after-body: scroll-anim.html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(plotly)
```

```{r}
#| label: load-data
# Read 10-year hourly weather telemetry for every recorded observation
files <- list.files("data", pattern = "\\.csv$", full.names = TRUE)

weather_hourly <- map(files, \(f) {
  read_csv(f, col_types = cols(.default = col_character()), show_col_types = FALSE)
}) |>
  list_rbind() |>
  filter(grepl(":53:", DATE)) |>
  mutate(
    datetime = as.POSIXct(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    dew_point = parse_number(HourlyDewPointTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind_speed = parse_number(HourlyWindSpeed),
    precip = replace_na(parse_number(na_if(HourlyPrecipitation, "T")), 0)
  ) |>
  filter(!is.na(datetime))

# Significant Florida natural disasters and storms during 2015-2024
disasters <- tribble(
  ~date, ~event,
  as.Date("2016-09-02"), "Hurricane Hermine",
  as.Date("2017-09-10"), "Hurricane Irma",
  as.Date("2018-10-10"), "Hurricane Michael",
  as.Date("2022-09-28"), "Hurricane Ian",
  as.Date("2022-12-24"), "Christmas Arctic Freeze",
  as.Date("2023-08-30"), "Hurricane Idalia",
  as.Date("2024-08-04"), "Hurricane Debby",
  as.Date("2024-09-26"), "Hurricane Helene",
  as.Date("2024-10-09"), "Hurricane Milton"
)

# Extract precise metric values during each disaster date for interactive hover markers
disaster_points <- map(1:nrow(disasters), \(i) {
  d <- disasters$date[i]
  ev <- disasters$event[i]
  sub <- weather_hourly |> filter(as.Date(datetime) == d)
  
  tibble(
    event = ev,
    date = d,
    datetime = sub$datetime[1],
    temp = if(grepl("Freeze", ev)) min(sub$temp, na.rm = TRUE) else max(sub$temp, na.rm = TRUE),
    dew_point = mean(sub$dew_point, na.rm = TRUE),
    pressure = min(sub$pressure, na.rm = TRUE),
    wind_speed = max(sub$wind_speed, na.rm = TRUE),
    precip = max(sub$precip, na.rm = TRUE)
  )
}) |> list_rbind()

# Vertical red dashed markers for natural disasters
disaster_shapes <- pmap(disaster_points, \(datetime, event, ...) {
  list(
    type = "line",
    x0 = datetime,
    x1 = datetime,
    y0 = 0,
    y1 = 1,
    yref = "paper",
    line = list(color = "#C62828", dash = "dash", width = 1.3)
  )
})
```

<div class="news-header">
<span class="news-badge">SPECIAL METEOROLOGICAL DISPATCH</span>
<h1 style="color: #ffffff; margin-top: 0.5rem; font-weight: 800;">A Decade Under the Storm: Inside Tampa Bay's 10-Year Weather Extremes</h1>
<p style="margin-bottom: 0.25rem;"><strong>By The Weather Desk</strong> | Tampa International Airport (KTPA) Telemetry Network</p>
<div class="news-station-tag">STATION: KTPA (72211012842) | LAT: 27.96° N | LON: -82.54° W | ELEV: 19 FT | PERIOD: 2015–2024</div>
</div>

<div class="records-grid">
<div class="record-card">
<div class="record-icon">🌡️</div>
<div class="record-value">98°F</div>
<div class="record-label">10-Year Record High</div>
<div class="record-event">Summer Heatwaves</div>
</div>
<div class="record-card">
<div class="record-icon">❄️</div>
<div class="record-value">29°F</div>
<div class="record-label">10-Year Record Low</div>
<div class="record-event">Dec 2022 Arctic Freeze</div>
</div>
<div class="record-card">
<div class="record-icon">🌪️</div>
<div class="record-value">28.69</div>
<div class="record-label">Lowest Pressure (inHg)</div>
<div class="record-event">Hurricane Irma Eye (2017)</div>
</div>
<div class="record-card">
<div class="record-icon">💨</div>
<div class="record-value">47 mph</div>
<div class="record-label">Max Sustained Wind</div>
<div class="record-event">Tropical Gale Peak</div>
</div>
<div class="record-card">
<div class="record-icon">🌧️</div>
<div class="record-value">16.14 in</div>
<div class="record-label">Wettest Single Day</div>
<div class="record-event">Hurricane Debby (2024)</div>
</div>
</div>

::: {.callout-note}
### Natural Disasters Tracked (Red Dashed Markers & Interactive Red Dots)
Across the decade, Tampa confronted historic tropical cyclones and extreme temperature anomalies. The red dashed vertical lines and interactive red markers on each graph highlight the 9 major events:

1. **Hurricane Hermine** (Sep 2, 2016) — Tropical gale & coastal flooding
2. **Hurricane Irma** (Sep 10, 2017) — Historic pressure drop (28.69 inHg), 46 mph sustained winds
3. **Hurricane Michael** (Oct 10, 2018) — Category 5 Gulf passage
4. **Hurricane Ian** (Sep 28, 2022) — Major storm surge threat, 36 mph winds
5. **Christmas Arctic Outbreak** (Dec 24, 2022) — Decade-record freeze down to 29°F
6. **Hurricane Idalia** (Aug 30, 2023) — Category 4 Gulf transit & high tide surge
7. **Hurricane Debby** (Aug 4, 2024) — Record 24-hr deluge (16.14 in)
8. **Hurricane Helene** (Sep 26, 2024) — Historic storm surge disaster, 39 mph winds
9. **Hurricane Milton** (Oct 9, 2024) — Direct Tampa Bay landfall, 16.00 in rain, 29.10 inHg pressure

*Tip: Hover over any red dot or dashed line on the charts to see the storm name and its exact telemetry reading!*
:::

## 1. Thermal Trends: Temperature Record

::::: {.news-chart-card}
:::: {.news-section-tag}
SURFACE CLIMATOLOGY BRIEFING
::::
### Heat Domes and Winter Frosts: A Decade of Temperature Shifts
<p class="news-lede">Tampa's subtropical marine climate maintains prolonged summer heat streaks where daily temperatures persistently top 85°F. Winter months bring brief, sharp cold snaps driven by continental polar air masses. The most severe cold outbreak occurred during the Christmas 2022 freeze, dropping temperatures to an extraordinary 29°F.</p>

<div class="stat-highlight-row">
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Peak Data Point</span>
<span class="stat-badge-value">98.0°F <small style="font-weight:normal; color:#64748b;">(Aug 11, 2023) | Daily Avg: 89.7°F</small></span>
</div>
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Average Year</span>
<span class="stat-badge-value">2023 <small style="font-weight:normal; color:#64748b;">(Annual Mean: 75.6°F)</small></span>
</div>
</div>

:::: {.scroll-graph}
```{r}
plot_ly() |>
  add_trace(data = weather_hourly, x = ~datetime, y = ~temp, type = "scattergl", mode = "lines",
            line = list(color = "#D84315", width = 1), name = "Hourly Temp") |>
  add_trace(data = disaster_points, x = ~datetime, y = ~temp, type = "scatter", mode = "markers",
            marker = list(color = "#C62828", size = 8, symbol = "circle"),
            name = "Disaster Event",
            text = ~paste0("<b>", event, "</b><br>Date: ", format(date, "%b %d, %Y"), "<br>Temp: ", temp, " °F"),
            hoverinfo = "text") |>
  layout(
    title = list(text = "Hourly Temperature (°F) — Every Recorded Observation (2015–2024)", x = 0),
    xaxis = list(title = "Timestamp", showgrid = TRUE),
    yaxis = list(title = "Temperature (°F)", showgrid = TRUE),
    shapes = disaster_shapes,
    margin = list(t = 40, b = 40, l = 50, r = 20)
  )
```
::::
:::::

## 2. Atmospheric Moisture: Dew Point Analysis

::::: {.news-chart-card}
:::: {.news-section-tag}
MOISTURE & HUMIDITY PROFILING
::::
### The Gulf Moisture Engine: Dew Points Above 75°F
<p class="news-lede">Dew point temperatures serve as meteorologists' premier measurement of true atmospheric moisture content. In Tampa, summer dew points frequently exceed 75°F—a threshold indicating tropical mugginess that fuels convective afternoon thunderstorms and supercharges landfalling tropical cyclones.</p>

<div class="stat-highlight-row">
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Peak Data Point</span>
<span class="stat-badge-value">82.0°F <small style="font-weight:normal; color:#64748b;">(Aug 7, 2024) | Daily Avg: 80.7°F</small></span>
</div>
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Average Year</span>
<span class="stat-badge-value">2015 <small style="font-weight:normal; color:#64748b;">(Annual Mean: 65.2°F)</small></span>
</div>
</div>

:::: {.scroll-graph}
```{r}
plot_ly() |>
  add_trace(data = weather_hourly, x = ~datetime, y = ~dew_point, type = "scattergl", mode = "lines",
            line = list(color = "#00695C", width = 1), name = "Hourly Dew Point") |>
  add_trace(data = disaster_points, x = ~datetime, y = ~dew_point, type = "scatter", mode = "markers",
            marker = list(color = "#C62828", size = 8, symbol = "circle"),
            name = "Disaster Event",
            text = ~paste0("<b>", event, "</b><br>Date: ", format(date, "%b %d, %Y"), "<br>Dew Point: ", round(dew_point, 1), " °F"),
            hoverinfo = "text") |>
  layout(
    title = list(text = "Hourly Dew Point (°F) — Every Recorded Observation (2015–2024)", x = 0),
    xaxis = list(title = "Timestamp", showgrid = TRUE),
    yaxis = list(title = "Dew Point (°F)", showgrid = TRUE),
    shapes = disaster_shapes,
    margin = list(t = 40, b = 40, l = 50, r = 20)
  )
```
::::
:::::

## 3. Barometric Pressure: Cyclone Footprints

::::: {.news-chart-card}
:::: {.news-section-tag}
SYNOPTIC BAROMETRY & HURRICANE TRACKING
::::
### The Fingerprint of Monster Storms: Dramatic Pressure Plunges
<p class="news-lede">Atmospheric pressure is the vital sign of tropical cyclones. Standard atmospheric pressure at KTPA rests around 30.00 to 30.20 inHg. On September 10–11, 2017, Hurricane Irma's core hammered Tampa, causing station pressure to plummet to a historic low of 28.69 inHg. In October 2024, Hurricane Milton triggered another collapse to 29.10 inHg as its eye passed near Tampa Bay.</p>

<div class="stat-highlight-row">
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Peak Data Point</span>
<span class="stat-badge-value">30.55 inHg <small style="font-weight:normal; color:#64748b;">(Jan 8, 2017) | Daily Avg: 30.52 inHg</small></span>
</div>
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Average Year</span>
<span class="stat-badge-value">2018 <small style="font-weight:normal; color:#64748b;">(Annual Mean: 30.06 inHg)</small></span>
</div>
</div>

:::: {.scroll-graph}
```{r}
plot_ly() |>
  add_trace(data = weather_hourly, x = ~datetime, y = ~pressure, type = "scattergl", mode = "lines",
            line = list(color = "#1565C0", width = 1), name = "Hourly Pressure") |>
  add_trace(data = disaster_points, x = ~datetime, y = ~pressure, type = "scatter", mode = "markers",
            marker = list(color = "#C62828", size = 8, symbol = "circle"),
            name = "Disaster Event",
            text = ~paste0("<b>", event, "</b><br>Date: ", format(date, "%b %d, %Y"), "<br>Pressure: ", pressure, " inHg"),
            hoverinfo = "text") |>
  layout(
    title = list(text = "Hourly Station Pressure (inHg) — Every Recorded Observation (2015–2024)", x = 0),
    xaxis = list(title = "Timestamp", showgrid = TRUE),
    yaxis = list(title = "Pressure (inHg)", showgrid = TRUE),
    shapes = disaster_shapes,
    margin = list(t = 40, b = 40, l = 50, r = 20)
  )
```
::::
:::::

## 4. Anemometer Telemetry: Wind Speeds & Tropical Gales

::::: {.news-chart-card}
:::: {.news-section-tag}
ANEMOMETER & GALE TELEMETRY
::::
### From Coastal Sea Breezes to Hurricane-Force Gales
<p class="news-lede">Average wind speeds in Tampa generally hover between 5 and 10 mph due to diurnal sea breezes. However, tropical cyclone passages dramatically spike wind velocities. Hurricane Irma (2017) recorded sustained winds reaching 46 mph at the airport, while Hurricane Helene (2024) generated sustained 39 mph gales driving unprecedented coastal storm surge across Tampa Bay.</p>

<div class="stat-highlight-row">
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Peak Data Point</span>
<span class="stat-badge-value">47.0 mph <small style="font-weight:normal; color:#64748b;">(Sep 10, 2017 - Irma) | Daily Avg: 24.4 mph</small></span>
</div>
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Average Year</span>
<span class="stat-badge-value">2020 <small style="font-weight:normal; color:#64748b;">(Annual Mean: 7.1 mph)</small></span>
</div>
</div>

:::: {.scroll-graph}
```{r}
plot_ly() |>
  add_trace(data = weather_hourly, x = ~datetime, y = ~wind_speed, type = "scattergl", mode = "lines",
            line = list(color = "#6A1B9A", width = 1), name = "Hourly Wind Speed") |>
  add_trace(data = disaster_points, x = ~datetime, y = ~wind_speed, type = "scatter", mode = "markers",
            marker = list(color = "#C62828", size = 8, symbol = "circle"),
            name = "Disaster Event",
            text = ~paste0("<b>", event, "</b><br>Date: ", format(date, "%b %d, %Y"), "<br>Wind Speed: ", wind_speed, " mph"),
            hoverinfo = "text") |>
  layout(
    title = list(text = "Hourly Wind Speed (mph) — Every Recorded Observation (2015–2024)", x = 0),
    xaxis = list(title = "Timestamp", showgrid = TRUE),
    yaxis = list(title = "Wind Speed (mph)", showgrid = TRUE),
    shapes = disaster_shapes,
    margin = list(t = 40, b = 40, l = 50, r = 20)
  )
```
::::
:::::

## 5. Hydrological Extremes: Precipitation Records

::::: {.news-chart-card}
:::: {.news-section-tag}
HYDROLOGICAL DISPATCH & RAINFALL TOTALS
::::
### The 2024 Deluges: Unprecedented Rain Events
<p class="news-lede">While Florida experiences intense daily summer rain bursts, 2024 produced catastrophic back-to-back precipitation extremes. On August 4, 2024, tropical feeder bands from Hurricane Debby inundated KTPA with 16.14 inches of rain in 24 hours—the wettest single day in the modern station record. Just two months later, Hurricane Milton dumped an additional 16.00 inches in a single evening.</p>

<div class="stat-highlight-row">
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Peak Data Point</span>
<span class="stat-badge-value">16.14 in <small style="font-weight:normal; color:#64748b;">(Aug 4, 2024 - Debby)</small></span>
</div>
<div class="stat-badge-item">
<span class="stat-badge-label">Highest Average Year</span>
<span class="stat-badge-value">2024 <small style="font-weight:normal; color:#64748b;">(Daily Avg: 0.48 in/day | Total: 175.1 in)</small></span>
</div>
</div>

:::: {.scroll-graph}
```{r}
plot_ly() |>
  add_trace(data = weather_hourly, x = ~datetime, y = ~precip, type = "scattergl", mode = "lines",
            line = list(color = "#0288D1", width = 1), name = "Hourly Precip") |>
  add_trace(data = disaster_points, x = ~datetime, y = ~precip, type = "scatter", mode = "markers",
            marker = list(color = "#C62828", size = 8, symbol = "circle"),
            name = "Disaster Event",
            text = ~paste0("<b>", event, "</b><br>Date: ", format(date, "%b %d, %Y"), "<br>Precipitation: ", precip, " in"),
            hoverinfo = "text") |>
  layout(
    title = list(text = "Hourly Precipitation (inches) — Every Recorded Observation (2015–2024)", x = 0),
    xaxis = list(title = "Timestamp", showgrid = TRUE),
    yaxis = list(title = "Precipitation (in)", showgrid = TRUE),
    shapes = disaster_shapes,
    margin = list(t = 40, b = 40, l = 50, r = 20)
  )
```
::::
:::::

## 6. Natural Disaster & Landfall Telemetry Dossier (2015–2024)

::::: {.news-chart-card}
:::: {.news-section-tag}
TROPICAL & EXTREME WEATHER INTELLIGENCE
::::
### Landfall & Storm Peak Telemetry Analysis
<p class="news-lede">When major hurricanes make landfall along the Florida coastline, Tampa International Airport (KTPA) records immediate atmospheric signatures—dramatic drops in station pressure, intense wind gusts, and torrential tropical rain bands. The comparative chart and telemetry dossier below highlight key meteorological data points captured <strong>right at landfall</strong> as well as each disaster's overall peak intensity at Tampa.</p>

:::: {.scroll-graph}
```{r}
#| label: disaster-comparison-chart
disaster_chart_data <- tribble(
  ~event, ~type, ~cat, ~landfall_time, ~landfall_loc, ~press_lf, ~wind_lf, ~temp_lf, ~min_press, ~max_wind, ~tot_rain,
  "Hermine", "Cyclone", "Cat 1", "Sep 2, 2016 01:30 EDT", "St. Marks, FL", 29.80, 26, 82, 29.74, 30, 6.70,
  "Irma", "Cyclone", "Cat 3/4", "Sep 10, 2017 15:35 EDT", "Marco Island, FL", 29.49, 25, 75, 28.69, 46, 12.45,
  "Michael", "Cyclone", "Cat 5", "Oct 10, 2018 13:00 EDT", "Mexico Beach, FL", 29.76, 21, 85, 29.72, 22, 0.54,
  "Ian", "Cyclone", "Cat 4", "Sep 28, 2022 15:05 EDT", "Cayo Costa, FL", 29.58, 23, 73, 29.53, 36, 4.19,
  "Xmas Freeze", "Freeze", "Historic", "Dec 24, 2022 07:00 EST", "Continental Polar", 30.28, 10, 32, 29.93, 22, 0.00,
  "Idalia", "Cyclone", "Cat 3", "Aug 30, 2023 07:45 EDT", "Keaton Beach, FL", 29.64, 23, 83, 29.52, 33, 6.89,
  "Debby", "Cyclone", "Cat 1", "Aug 4, 2024 07:00 EDT", "Steinhatchee, FL", 29.82, 7, 79, 29.67, 34, 17.68,
  "Helene", "Cyclone", "Cat 4", "Sep 26, 2024 23:10 EDT", "Perry, FL", 29.50, 24, 82, 29.29, 39, 4.77,
  "Milton", "Cyclone", "Cat 3", "Oct 9, 2024 20:30 EDT", "Siesta Key, FL", 29.12, 33, 71, 29.12, 33, 16.03
)

plot_ly(disaster_chart_data, x = ~event) |>
  add_bars(
    y = ~max_wind,
    name = "Peak Sustained Wind (mph)",
    marker = list(color = "#7B1FA2"),
    text = ~paste0("<b>", event, "</b> (", cat, ")<br>Landfall: ", landfall_time, "<br>Location: ", landfall_loc, "<br><b>At Landfall:</b><br>• Wind: ", wind_lf, " mph<br>• Pressure: ", press_lf, " inHg<br>• Temp: ", temp_lf, "°F<br><b>Peak Wind:</b> ", max_wind, " mph"),
    hoverinfo = "text"
  ) |>
  add_bars(
    y = ~tot_rain,
    name = "Storm Deluge Rain (in)",
    marker = list(color = "#0288D1"),
    text = ~paste0("<b>", event, "</b><br>Storm Rain Total: ", tot_rain, " in<br>At Landfall Temp: ", temp_lf, "°F"),
    hoverinfo = "text"
  ) |>
  add_lines(
    y = ~min_press,
    name = "Min Pressure (inHg)",
    yaxis = "y2",
    line = list(color = "#C62828", width = 3),
    text = ~paste0("<b>", event, "</b><br>Lowest Pressure: ", min_press, " inHg<br>At Landfall: ", press_lf, " inHg"),
    hoverinfo = "text"
  ) |>
  layout(
    title = list(text = "10-Year Natural Disaster Telemetry & Landfall Comparison", x = 0),
    xaxis = list(title = "Natural Disaster Event"),
    yaxis = list(title = "Wind (mph) & Rainfall (inches)", showgrid = TRUE),
    yaxis2 = list(title = "Lowest Pressure (inHg)", overlaying = "y", side = "right", range = c(28.2, 30.8), showgrid = FALSE),
    barmode = "group",
    legend = list(orientation = "h", x = 0.1, y = 1.15),
    margin = list(t = 60, b = 40, l = 50, r = 50)
  )
```
::::

<div class="disaster-table-container">
<table class="disaster-table">
<thead>
<tr>
<th>Event & Category</th>
<th>Official Landfall / Peak</th>
<th>Landfall Location</th>
<th>Data Points Right at Landfall (KTPA)</th>
<th>Storm Peak Extremes (KTPA)</th>
<th>Primary Tampa Bay Impact</th>
</tr>
</thead>
<tbody>
<tr>
<td><span class="disaster-badge-cyclone">Cat 1 Hurricane</span><br><strong>Hermine</strong></td>
<td>Sep 2, 2016<br>01:30 EDT</td>
<td>St. Marks / Big Bend, FL</td>
<td><span class="telemetry-val">82°F</span> | <span class="telemetry-val">29.80 inHg</span> | <span class="telemetry-val">26 mph</span></td>
<td>Pressure: <span class="telemetry-val">29.74 inHg</span><br>Peak Wind: <span class="telemetry-val">30 mph</span><br>Rain: <span class="telemetry-val">6.70 in</span></td>
<td>Outer tropical feeder bands brought coastal road flooding and tidal surges across the Bay.</td>
</tr>
<tr>
<td><span class="disaster-badge-cyclone">Cat 3/4 Hurricane</span><br><strong>Irma</strong></td>
<td>Sep 10, 2017<br>15:35 EDT</td>
<td>Marco Island, FL<br><em>(Eye passage: Sep 11 02:00 EDT)</em></td>
<td><span class="telemetry-val">75°F</span> | <span class="telemetry-val">29.49 inHg</span> | <span class="telemetry-val">25 mph</span><br><em>(At Eye: 28.69 inHg, 46 mph)</em></td>
<td>Pressure: <span class="telemetry-val">28.69 inHg (Decade Low)</span><br>Peak Wind: <span class="telemetry-val">46 mph (66 mph gust)</span><br>Rain: <span class="telemetry-val">12.45 in</span></td>
<td>Core battered Tampa with 66 mph gusts, record barometric plunge to 28.69 inHg, and widespread blackouts.</td>
</tr>
<tr>
<td><span class="disaster-badge-cyclone">Cat 5 Hurricane</span><br><strong>Michael</strong></td>
<td>Oct 10, 2018<br>13:00 EDT</td>
<td>Mexico Beach, FL</td>
<td><span class="telemetry-val">85°F</span> | <span class="telemetry-val">29.76 inHg</span> | <span class="telemetry-val">21 mph</span></td>
<td>Pressure: <span class="telemetry-val">29.72 inHg</span><br>Peak Wind: <span class="telemetry-val">22 mph</span><br>Rain: <span class="telemetry-val">0.54 in</span></td>
<td>Intense offshore trajectory drained shallow bays through reverse storm surge before outer squalls hit.</td>
</tr>
<tr>
<td><span class="disaster-badge-cyclone">Cat 4 Hurricane</span><br><strong>Ian</strong></td>
<td>Sep 28, 2022<br>15:05 EDT</td>
<td>Cayo Costa / Fort Myers, FL</td>
<td><span class="telemetry-val">73°F</span> | <span class="telemetry-val">29.58 inHg</span> | <span class="telemetry-val">23 mph</span></td>
<td>Pressure: <span class="telemetry-val">29.53 inHg</span><br>Peak Wind: <span class="telemetry-val">36 mph</span><br>Rain: <span class="telemetry-val">4.19 in</span></td>
<td>Massive negative surge drained Hillsborough Bay dry for hours before tropical gale winds returned water.</td>
</tr>
<tr>
<td><span class="disaster-badge-freeze">Historic Cold Wave</span><br><strong>Christmas Freeze</strong></td>
<td>Dec 24, 2022<br>07:00 EST</td>
<td>Continental Polar Air Outbreak</td>
<td><span class="telemetry-val">32°F</span> | <span class="telemetry-val">30.28 inHg</span> | <span class="telemetry-val">10 mph</span></td>
<td>Min Temp: <span class="telemetry-val">29.0°F (Decade Low)</span><br>Dew Point: <span class="telemetry-val">18.0°F</span><br>Max Pressure: <span class="telemetry-val">30.28 inHg</span></td>
<td>Severe 3-day hard freeze across Central Florida damaging subtropical agriculture and bursting water pipes.</td>
</tr>
<tr>
<td><span class="disaster-badge-cyclone">Cat 3 Hurricane</span><br><strong>Idalia</strong></td>
<td>Aug 30, 2023<br>07:45 EDT</td>
<td>Keaton Beach, FL</td>
<td><span class="telemetry-val">83°F</span> | <span class="telemetry-val">29.64 inHg</span> | <span class="telemetry-val">23 mph</span></td>
<td>Pressure: <span class="telemetry-val">29.52 inHg</span><br>Peak Wind: <span class="telemetry-val">33 mph</span><br>Rain: <span class="telemetry-val">6.89 in</span></td>
<td>Coincided with astronomical king tides, pushing 4–5 ft saltwater inundation into Bayshore Blvd.</td>
</tr>
<tr>
<td><span class="disaster-badge-cyclone">Cat 1 Deluge</span><br><strong>Debby</strong></td>
<td>Aug 4, 2024<br>07:00 EDT</td>
<td>Steinhatchee, FL</td>
<td><span class="telemetry-val">79°F</span> | <span class="telemetry-val">29.82 inHg</span> | <span class="telemetry-val">7 mph</span></td>
<td>Pressure: <span class="telemetry-val">29.67 inHg</span><br>Peak Wind: <span class="telemetry-val">34 mph</span><br>Rain: <span class="telemetry-val">17.68 in (16.14 in/24hr)</span></td>
<td>Historic 24-hr rainfall record (16.14 in) causing historic flash flooding across Hillsborough County.</td>
</tr>
<tr>
<td><span class="disaster-badge-cyclone">Cat 4 Surge</span><br><strong>Helene</strong></td>
<td>Sep 26, 2024<br>23:10 EDT</td>
<td>Perry / Big Bend, FL</td>
<td><span class="telemetry-val">82°F</span> | <span class="telemetry-val">29.50 inHg</span> | <span class="telemetry-val">24 mph</span></td>
<td>Pressure: <span class="telemetry-val">29.29 inHg</span><br>Peak Wind: <span class="telemetry-val">39 mph (55+ mph gust)</span><br>Rain: <span class="telemetry-val">4.77 in</span></td>
<td>Catastrophic 6–7 ft storm surge devastated coastal Pinellas & Hillsborough barrier islands.</td>
</tr>
<tr>
<td><span class="disaster-badge-cyclone">Cat 3 Eyewall</span><br><strong>Milton</strong></td>
<td>Oct 9, 2024<br>20:30 EDT</td>
<td>Siesta Key / Sarasota, FL<br><em>(50 miles S of KTPA)</em></td>
<td><span class="telemetry-val">71°F</span> | <span class="telemetry-val">29.12 inHg</span> | <span class="telemetry-val">33 mph</span></td>
<td>Pressure: <span class="telemetry-val">29.12 inHg (Eyewall Dip)</span><br>Peak Wind: <span class="telemetry-val">33 mph (50+ mph gust)</span><br>Rain: <span class="telemetry-val">16.03 in</span></td>
<td>Direct northern eyewall transit dumped 16 inches of torrential rain with fierce squalls and building damage.</td>
</tr>
</tbody>
</table>
</div>
:::::

