---
title: "The Tampa Bay Climate Chronicle"
subtitle: "A 10-Year Meteorological Investigation & Extreme Weather Dispatch (2015–2024)"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    css: styles.css
    include-after-body: scroll-anim.html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(plotly)
```

```{r}
#| label: load-data
# Read 10-year hourly weather telemetry and aggregate daily statistics
files <- list.files("data", pattern = "\\.csv$", full.names = TRUE)

weather <- map(files, \(f) {
  read_csv(f, col_types = cols(.default = col_character()), show_col_types = FALSE)
}) |>
  list_rbind() |>
  mutate(
    date = as.Date(DATE),
    temp = parse_number(HourlyDryBulbTemperature),
    dew_point = parse_number(HourlyDewPointTemperature),
    pressure = parse_number(HourlyStationPressure),
    wind_speed = parse_number(HourlyWindSpeed),
    precip = parse_number(na_if(HourlyPrecipitation, "T"))
  ) |>
  mutate(precip = replace_na(precip, 0)) |>
  group_by(date) |>
  summarize(
    temp = mean(temp, na.rm = TRUE),
    dew_point = mean(dew_point, na.rm = TRUE),
    pressure = mean(pressure, na.rm = TRUE),
    wind_speed = mean(wind_speed, na.rm = TRUE),
    precip = sum(precip, na.rm = TRUE),
    .groups = "drop"
  )

# Significant Florida natural disasters and storms during 2015-2024
disasters <- tribble(
  ~date, ~event,
  as.Date("2016-09-02"), "Hurricane Hermine",
  as.Date("2017-09-10"), "Hurricane Irma",
  as.Date("2018-10-10"), "Hurricane Michael",
  as.Date("2022-09-28"), "Hurricane Ian",
  as.Date("2022-12-24"), "Christmas Arctic Freeze",
  as.Date("2023-08-30"), "Hurricane Idalia",
  as.Date("2024-08-04"), "Hurricane Debby",
  as.Date("2024-09-26"), "Hurricane Helene",
  as.Date("2024-10-09"), "Hurricane Milton"
)

weather <- left_join(weather, disasters, by = "date")
```

<div class="news-header">
  <span class="news-badge">SPECIAL METEOROLOGICAL DISPATCH</span>
  <h1 style="color: #ffffff; margin-top: 0.5rem; font-weight: 800;">A Decade Under the Storm: Inside Tampa Bay's 10-Year Weather Extremes</h1>
  <p style="margin-bottom: 0.25rem;"><strong>By The Weather Desk</strong> | Tampa International Airport (KTPA) Telemetry Network</p>
  <div class="news-station-tag">STATION: KTPA (72211012842) | LAT: 27.96° N | LON: -82.54° W | ELEV: 19 FT | PERIOD: 2015–2024</div>
</div>

<div class="records-grid">
  <div class="record-card">
    <div class="record-icon">🌡️</div>
    <div class="record-value">98°F</div>
    <div class="record-label">10-Year Record High</div>
    <div class="record-event">Summer Heatwaves</div>
  </div>
  <div class="record-card">
    <div class="record-icon">❄️</div>
    <div class="record-value">29°F</div>
    <div class="record-label">10-Year Record Low</div>
    <div class="record-event">Dec 2022 Arctic Freeze</div>
  </div>
  <div class="record-card">
    <div class="record-icon">🌪️</div>
    <div class="record-value">28.69</div>
    <div class="record-label">Lowest Pressure (inHg)</div>
    <div class="record-event">Hurricane Irma Eye (2017)</div>
  </div>
  <div class="record-card">
    <div class="record-icon">💨</div>
    <div class="record-value">47 mph</div>
    <div class="record-label">Max Sustained Wind</div>
    <div class="record-event">Tropical Gale Peak</div>
  </div>
  <div class="record-card">
    <div class="record-icon">🌧️</div>
    <div class="record-value">16.14 in</div>
    <div class="record-label">Wettest Single Day</div>
    <div class="record-event">Hurricane Debby (2024)</div>
  </div>
</div>

::: {.callout-note}
### Natural Disasters Tracked (Red Dashed Markers)
Across the decade, Tampa confronted historic tropical cyclones and extreme temperature anomalies. The red dashed lines and highlighted dots mark landfall and close-passage dates for **Hurricane Hermine (2016)**, **Hurricane Irma (2017)**, **Hurricane Michael (2018)**, **Hurricane Ian (2022)**, the **Christmas 2022 Arctic Outbreak**, **Hurricane Idalia (2023)**, **Hurricane Debby (2024)**, **Hurricane Helene (2024)**, and **Hurricane Milton (2024)**. As you scroll down, each interactive chart graphs itself dynamically across the 10-year timeline!
:::

## 1. Thermal Trends: Temperature Record

::::: {.news-chart-card}
:::: {.news-section-tag}
SURFACE CLIMATOLOGY BRIEFING
::::
### Heat Domes and Winter Frosts: A Decade of Temperature Shifts
Tampa's subtropical marine climate maintains prolonged summer heat streaks where daily averages persistently top 85°F. Winter months bring brief, sharp cold snaps driven by continental polar air masses. The most severe cold outbreak occurred during the Christmas 2022 freeze, dropping temperatures to an extraordinary 29°F.

:::: {.scroll-graph}
```{r}
p_temp <- ggplot(weather, aes(x = date, y = temp)) +
  geom_area(fill = "#FF8A65", alpha = 0.25) +
  geom_line(color = "#D84315", linewidth = 0.6) +
  geom_vline(data = disasters, aes(xintercept = date), linetype = "dashed", color = "#C62828", alpha = 0.4) +
  geom_point(data = filter(weather, !is.na(event)), 
             aes(text = paste0("<b>", event, "</b><br>Date: ", date, "<br>Daily Avg Temp: ", round(temp, 1), " °F")),
             color = "#C62828", size = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0.01, 0)) +
  labs(title = "Daily Average Temperature (°F) — Tampa International Airport (2015–2024)", x = "Year", y = "Temperature (°F)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

suppressWarnings(ggplotly(p_temp, tooltip = c("x", "y", "text")))
```
::::
:::::

## 2. Atmospheric Moisture: Dew Point Analysis

::::: {.news-chart-card}
:::: {.news-section-tag}
MOISTURE & HUMIDITY PROFILING
::::
### The Gulf Moisture Engine: Dew Points Above 75°F
Dew point temperatures serve as meteorologists' premier measurement of true atmospheric moisture content. In Tampa, summer dew points frequently exceed 75°F—a threshold indicating tropical mugginess that fuels convective afternoon thunderstorms and supercharges landfalling tropical cyclones.

:::: {.scroll-graph}
```{r}
p_dew <- ggplot(weather, aes(x = date, y = dew_point)) +
  geom_area(fill = "#80CBC4", alpha = 0.25) +
  geom_line(color = "#00695C", linewidth = 0.6) +
  geom_vline(data = disasters, aes(xintercept = date), linetype = "dashed", color = "#C62828", alpha = 0.4) +
  geom_point(data = filter(weather, !is.na(event)), 
             aes(text = paste0("<b>", event, "</b><br>Date: ", date, "<br>Daily Avg Dew Point: ", round(dew_point, 1), " °F")),
             color = "#C62828", size = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0.01, 0)) +
  labs(title = "Daily Average Dew Point (°F) — Atmospheric Moisture Index (2015–2024)", x = "Year", y = "Dew Point (°F)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

suppressWarnings(ggplotly(p_dew, tooltip = c("x", "y", "text")))
```
::::
:::::

## 3. Barometric Pressure: Cyclone Footprints

::::: {.news-chart-card}
:::: {.news-section-tag}
SYNOPTIC BAROMETRY & HURRICANE TRACKING
::::
### The Fingerprint of Monster Storms: Dramatic Pressure Plunges
Atmospheric pressure is the vital sign of tropical cyclones. Standard atmospheric pressure at KTPA rests around 30.00 to 30.20 inHg. On September 10–11, 2017, Hurricane Irma's core hammered Tampa, causing station pressure to plummet to a historic low of 28.69 inHg. In October 2024, Hurricane Milton triggered another collapse to 29.10 inHg as its eye passed near Tampa Bay.

:::: {.scroll-graph}
```{r}
p_press <- ggplot(weather, aes(x = date, y = pressure)) +
  geom_area(fill = "#90CAF9", alpha = 0.25) +
  geom_line(color = "#1565C0", linewidth = 0.6) +
  geom_vline(data = disasters, aes(xintercept = date), linetype = "dashed", color = "#C62828", alpha = 0.4) +
  geom_point(data = filter(weather, !is.na(event)), 
             aes(text = paste0("<b>", event, "</b><br>Date: ", date, "<br>Station Pressure: ", round(pressure, 2), " inHg")),
             color = "#C62828", size = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0.01, 0)) +
  labs(title = "Daily Average Station Pressure (inHg) — Cyclone Signatures (2015–2024)", x = "Year", y = "Station Pressure (inHg)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

suppressWarnings(ggplotly(p_press, tooltip = c("x", "y", "text")))
```
::::
:::::

## 4. Anemometer Telemetry: Wind Speeds & Tropical Gales

::::: {.news-chart-card}
:::: {.news-section-tag}
ANEMOMETER & GALE TELEMETRY
::::
### From Coastal Sea Breezes to Hurricane-Force Gales
Daily average wind speeds in Tampa generally hover between 5 and 10 mph due to diurnal sea breezes. However, tropical cyclone passages dramatically spike wind velocities. Hurricane Irma (2017) recorded sustained winds reaching 46 mph at the airport, while Hurricane Helene (2024) generated sustained 39 mph gales driving unprecedented coastal storm surge across Tampa Bay.

:::: {.scroll-graph}
```{r}
p_wind <- ggplot(weather, aes(x = date, y = wind_speed)) +
  geom_area(fill = "#CE93D8", alpha = 0.25) +
  geom_line(color = "#6A1B9A", linewidth = 0.6) +
  geom_vline(data = disasters, aes(xintercept = date), linetype = "dashed", color = "#C62828", alpha = 0.4) +
  geom_point(data = filter(weather, !is.na(event)), 
             aes(text = paste0("<b>", event, "</b><br>Date: ", date, "<br>Daily Avg Wind Speed: ", round(wind_speed, 1), " mph")),
             color = "#C62828", size = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0.01, 0)) +
  labs(title = "Daily Average Wind Speed (mph) — 10-Year Anemometer Log (2015–2024)", x = "Year", y = "Wind Speed (mph)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

suppressWarnings(ggplotly(p_wind, tooltip = c("x", "y", "text")))
```
::::
:::::

## 5. Hydrological Extremes: Precipitation Records

::::: {.news-chart-card}
:::: {.news-section-tag}
HYDROLOGICAL DISPATCH & RAINFALL TOTALS
::::
### The 2024 Deluges: Unprecedented Rain Events
While Florida experiences intense daily summer rain bursts, 2024 produced catastrophic back-to-back precipitation extremes. On August 4, 2024, tropical feeder bands from Hurricane Debby inundated KTPA with 16.14 inches of rain in 24 hours—the wettest single day in the modern station record. Just two months later, Hurricane Milton dumped an additional 16.00 inches in a single evening.

:::: {.scroll-graph}
```{r}
p_precip <- ggplot(weather, aes(x = date, y = precip)) +
  geom_col(fill = "#0288D1", width = 1) +
  geom_vline(data = disasters, aes(xintercept = date), linetype = "dashed", color = "#C62828", alpha = 0.4) +
  geom_point(data = filter(weather, !is.na(event)), 
             aes(text = paste0("<b>", event, "</b><br>Date: ", date, "<br>Precipitation: ", round(precip, 2), " in")),
             color = "#C62828", size = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0.01, 0)) +
  labs(title = "Daily Total Precipitation (inches) — 10-Year Rainfall Log (2015–2024)", x = "Year", y = "Precipitation (in)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())

suppressWarnings(ggplotly(p_precip, tooltip = c("x", "y", "text")))
```
::::
:::::
